---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# labno

<!-- badges: start -->
<!-- badges: end -->

`labno` is a package that makes it easier to work with sample identifiers used
at the North West Genomic Laboratory Hub.

Each sample is given a DNA "lab number" which is stored as "labno" in sample databases.

## Installation

You can install the development version of labno from [GitHub](https://github.com/) with:

``` r
# install.packages("pak")
pak::pak("joe-m-shaw/labno")
```

## Information Governance

**No patient identifiable information should be included in this repository.**

If you spot some, please let me know.

For the purpose of testing functions, I have used generic examples for 
worksheet (WS123456) and lab number (12345678) values.

Where examples of patient names are required, I have used character names from
novels by Leo Tolstoy (Anna Karenina, Pierre Bezukhov etc).

## Example

You can use `labno` to extract sample identifiers stored in filenames, which is useful when doing data projects involving multiple sample files.

A more detailed example is given in the "Reading files" vignette.

```{r}

library(labno)

data <- data.frame(
  "filename" = c("WS123456_12345678a_AnnaKarenina"),
  "gene_result" = c(1)
)

```

Using `labno`, you can quickly extract the identifiers from the filename column.

```{r}

mutate_ids(data)

```