---
title: "Reading files"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{reading-files}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(labno)
library(readr)
library(purrr)
library(ggplot2)
```

Here is an common scenario of how to use `labno`.

There are multiple files with results in, and the sample identifiers are only
in the filenames.

```{r files}

files <- list.files(path = "data/",
           pattern = ".*.csv",
           full.names = TRUE)

files

```

Each file consists of results for different genes.

I have randomly generated a "score" for each gene. This is not real clinical data.

```{r example-file, echo=FALSE}

file1 <- read_csv(file = files[1],
                  show_col_types = FALSE)

file1

```

We create a simple function by combining `read_csv` from `readr` with the
`mutate_ids` function from `labno`.

```{r function}

read_csv_with_ids <- function(file) {
  
  output <- read_csv(file,
                     show_col_types = FALSE) |> 
    mutate_ids(file)
  
  return(output)
  
}

```

Using `map` from `purrr` we can iterate this function over our file list and 
bind the data together.

```{r read-data}

all_data <- files |> 
  map(\(files) read_csv_with_ids(files)) |> 
  list_rbind()

```

This gives us the data in a single dataframe with each row annotated with the
correct sample identifiers.

```{r data, echo = FALSE}

knitr::kable(all_data)

```

Then we can explore this collated data using `ggplot2`.

```{r plots, echo = FALSE, width = 6}

gene_plot <- ggplot(all_data, aes(x = gene, y = score)) +
  geom_point(shape = 21) +
  theme_bw()+
  theme(axis.text.x = element_text(angle = 90)) +
  facet_wrap(~name)

gene_plot
  
```
